### 2014-10-12 writing/AAAI SSS '15/abstract.tex
Adding results summary, cutting an Icard paper.
Edits

\begin{abstract}
  Natural logic offers a powerful relational conception of meaning
  that is a natural counterpart to distributed semantic
  representations, which have proven valuable in a wide range of
  sophisticated language tasks. However, it remains an open question
  whether it is possible to train distributed representations to
  support the rich, diverse logical reasoning captured by natural
  logic. We address this question a simple neural network-based
  model for learning embeddings.  Our experiments evaluate the models' 
  ability to learn the basic algebra of natural logic relations from 
  simulated data and from the WordNet noun graph.  We find that the model
  can learn to do inference over relations with near perfect precision
  from moderate amounts of training data.
\end{abstract}

>>>>

\begin{abstract}
  Natural logic offers a powerful relational conception of meaning
  that is a natural counterpart to distributed semantic
  representations, which have proven valuable in a wide range of
  sophisticated language tasks. However, it remains an open question
  whether it is possible to train distributed representations to
  support the rich, diverse logical reasoning captured by natural
  logic. We address this question using two neural network-based
  models for learning embeddings: plain neural networks and neural
  tensor networks. Our experiments evaluate the models' ability to
  learn the basic algebra of natural logic relations from simulated
  data and from the WordNet noun graph.  The overall positive results
  are promising for the future of learned distributed representations in
  the applied modeling of logical semantics.
\end{abstract}

### 2014-10-12 writing/AAAI SSS '15/intro.tex
Edits

Natural logic aligns well with distributed (e.g., vector)
representations, which also naturally model meaning relationally.
Distributed representations have been used successfully in a wide
array of sophisticated language tasks \cite{collobert2011natural}. %,
% including sentiment analysis, analogy completion, relation
% extraction, and named entity recognition.
However, it remains an open question whether it is possible to train
such representations to support the rich, diverse logical reasoning
captured by natural logic; while they excel at
synonymy (similarity), the results are more mixed for entailment,
contradiction, and mutual consistency.  Using the natural logic of
\cite{maccartney2009extended} as our formal model, we address this
open question for a simple neural network-based classification model. 
The natural logic is built from the seven relations defined in
Table~\ref{b-table}. Its formal properties are now well-understood
\cite{Icard:Moss:2013:LILT}, so it provides a rigorous
set of goals for our neural models. To keep the discussion manageable,
we limit attention to experiments involving the lexicon; for a more
extended treatment of complex expressions involving logical
connectives and quantifiers, see \citet{Bowman:Potts:Manning:2014}.

In our experiments, we evaluate these models' ability to learn the
basic algebra of natural logic relations from simulated data and from
the WordNet noun graph. The simulated data help us to achieve analytic
insights into what the models learn, and WordNet data show how they
fare in the real world.  We find that only the NTN is able to fully
learn the underlying algebra, but that both models excel in the 
WordNet experiment.

>>>>

Natural logic aligns well with distributed (e.g., vector)
representations, which also naturally model meaning relationally.
Distributed representations have been used successfully in a wide
array of sophisticated language tasks (e.g., \cite{collobert2011natural}). %,
% including sentiment analysis, analogy completion, relation
% extraction, and named entity recognition.
However, it remains an open question whether it is possible to train
such representations to support the rich, diverse logical reasoning
captured by natural logic; while they excel at
synonymy (similarity), the results are more mixed for entailment,
contradiction, and mutual consistency.  Using the natural logic of
\cite{maccartney2009extended} as our formal model, we address this
open question for a simple neural network-based classification model. 
The natural logic is built from the seven relations defined in
Table~\ref{b-table}. Its formal properties are now well-understood
\cite{Icard:Moss:2013:LILT}, so it provides a rigorous
set of goals for our neural models. To keep the discussion manageable,
we limit attention to experiments involving the lexicon; for a more
extended treatment of complex expressions involving logical
connectives and quantifiers, see \citet{Bowman:Potts:Manning:2014}.

In our experiments, we evaluate these models' ability to learn the
basic algebra of natural logic relations from simulated data and from
the WordNet noun graph. The simulated data help us to achieve analytic
insights into what the models learn, and the WordNet data show how they
fare with a real natural language vocabulary.  We find that only the NTN is able to fully
learn the underlying algebra, but that both models excel in the 
WordNet experiment.

### 2014-10-12 writing/AAAI SSS '15/join.tex
Edits

\paragraph{Experiments}
To test our models' ability to learn these inferential patterns, we
create artificial boolean structures in which terms denote sets of entities from
a small domain (e.g., Figure~\ref{lattice-figure}), employ logical
deduction to identify the valid statements, divide those into train
and test sets, and remove from the test set those statements which
cannot be proven from the training statements
(Figure~\ref{unprovable}).
% using the logic depicted in Figure~\ref{lattice-figure}.
%
%
% We test the model's ability to learn this behavior by creating
% artificial data sets of terms which represent sets of numbers. Since
% MacCartney and Manning's set of relations hold between sets as well as
% between sentences, we can use the underlying set structure to generate
% the all of the relations that hold between any pair of these terms, as
% in Figure \ref{lattice-figure}. We train the model defined above on a
% subset of these relations, but rather then presenting the model with a
% pair of tree-structured sentences as inputs, simply present it with
% two single terms, each of which corresponds to a single vector in the
% (randomly initialized) vocabulary matrix $V$, ensuring that the model
% has no information about the terms being compared except the relations
% between them.
%
In our experiments, we create 80 randomly generated sets drawn from a
domain of seven elements. This yields 6400 statements about pairs of
formulae, of which 3200 are chosen as a test set, and that test set is
further reduced to the 2960 examples that can be provably derived from
the training data. We trained both the NN model and the NTN model on
these data
sets% . In both cases, the models were implemented exactly as
% described in Section~\ref{methods}, but since the items being compared
% are single terms rather than full tree structures, the composition
% layer was not involved, and the two models differed only in which
% layer function was used for the comparison layer. We simply present
% the models with two single terms, each of which corresponds to a
% single vector in the (randomly initialized) vocabulary matrix $V$,
% thereby ensuring that the model has no information about the terms
% being compared except the relations between them.

\paragraph{Results} 
We found that the NTN model worked best with 11-dimensional vector
representations for the 80 sets and a 90-dimensional feature vector
for the classifier, though the performance of the model was not highly
dependant on either dimensionality setting. 
Averaging over five randomly generated datasets, the model was able to correctly label 98.1\% (standard deviation $\sigma = 1.5\%$) of the provably derivable test examples, and 87.7\%
($\sigma = 8.0\%$) of the remaining test examples. The simpler NN worked
best with 11 and 75 dimensions, respectively, but was able to achieve
accuracies of only 84.8\% ($\sigma = 1.9\%$) and 68.7\% ($\sigma = 3.5\%$),
respectively. Training accuracy was 99.8\% ($\sigma = 0.1\%$) for the NTN and 94.7\% ($\sigma = 2.0\%$) for
the NN.

>>>>

\paragraph{Experiments}
To test our models' ability to learn these inferential patterns, we
create artificial boolean structures in which terms denote sets of entities from
a small domain (e.g., Figure~\ref{lattice-figure}), employ logical
deduction to identify the valid statements, divide those into train
and test sets, and remove from the test set those statements which
cannot be proven from the training statements
(Figure~\ref{unprovable}).
% using the logic depicted in Figure~\ref{lattice-figure}.
%
%
% We test the model's ability to learn this behavior by creating
% artificial data sets of terms which represent sets of numbers. Since
% MacCartney and Manning's set of relations hold between sets as well as
% between sentences, we can use the underlying set structure to generate
% the all of the relations that hold between any pair of these terms, as
% in Figure \ref{lattice-figure}. We train the model defined above on a
% subset of these relations, but rather then presenting the model with a
% pair of tree-structured sentences as inputs, simply present it with
% two single terms, each of which corresponds to a single vector in the
% (randomly initialized) vocabulary matrix $V$, ensuring that the model
% has no information about the terms being compared except the relations
% between them.
%
In our experiments, we create 80 randomly generated sets drawn from a
domain of seven elements. This yields 6400 statements about pairs of
formulae, of which 3200 are chosen as a test set, and that test set is
further reduced to the 2960 examples that can be provably derived from
the training data. We trained both the NN model and the NTN model on
these data
sets.% . In both cases, the models were implemented exactly as
% described in Section~\ref{methods}, but since the items being compared
% are single terms rather than full tree structures, the composition
% layer was not involved, and the two models differed only in which
% layer function was used for the comparison layer. We simply present
% the models with two single terms, each of which corresponds to a
% single vector in the (randomly initialized) vocabulary matrix $V$,
% thereby ensuring that the model has no information about the terms
% being compared except the relations between them.

\paragraph{Results} 
We found that the NTN model worked best with 11-dimensional vector
representations for the 80 sets and a 90-dimensional feature vector
for the classifier, though the performance of the model was not highly
dependant on either dimensionality setting. 
Averaging over five randomly generated data sets, the model was able to correctly label 98.1\% (standard deviation $\sigma = 1.5\%$) of the provably derivable test examples, and 87.7\%
($\sigma = 8.0\%$) of the remaining test examples. The simpler NN worked
best with 11 and 75 dimensions, respectively, but was able to achieve
accuracies of only 84.8\% ($\sigma = 1.9\%$) and 68.7\% ($\sigma = 3.5\%$),
respectively. Training accuracy was 99.8\% ($\sigma = 0.1\%$) for the NTN and 94.7\% ($\sigma = 2.0\%$) for
the NN.

### 2014-10-12 writing/AAAI SSS '15/methods.tex
Edits

We build an embedding-based model using the method of
\cite{Bowman:Potts:Manning:2014}, which is centered on the task of
labeling a pair of words or sentences with one of a small set of
logical relations. The architecture of the model that we use, which is
limited to only pairs of single terms (such as words), is depicted
in Figure~\ref{sample-figure}. The model represents the two input
terms as embeddings, which are fed into a comparison function based
on one of two types of neural network layer functions to produce a
representation for the relationship between the two terms. This
representation is then fed into a softmax classifier, which outputs a
distribution over possible labels. The entire network, including the
embeddings, are trained using backpropagation.

The simpler version of the comparison simply concatenates the two
input vectors before feeding them into a standard neural network (NN)
layer.  The more powerful neural tensor network (NTN) version uses an
additional third-order tensor parameter to allow for multiplicative
interactions between the two inputs \cite{chen2013learning}. For more
details on the implementation and training of the layer functions, see
\cite{Bowman:Potts:Manning:2014}.

@@

This model used here differs from the one described in that work in
two ways. Because the inputs are single terms, we do not use
a composition function here. Also, for our experiment on WordNet data, 
we introduce an additional neural network layer between the embedding 
input and the comparison function, which facilitates initializing the 
embeddings themselves from pretrained vectors and was found to help 
performance in that setting.

>>>>

We build embedding-based models using the method of
\cite{Bowman:Potts:Manning:2014}, which is centered on the task of
labeling a pair of words or sentences with one of a small set of
logical relations. The architecture that we use, which is limited to
only pairs of single terms (such as words), is depicted in
Figure~\ref{sample-figure}. The model represents the two input terms
as embeddings, which are fed into a comparison function based on one
of two types of neural network layer functions to produce a
representation for the relationship between the two terms. This
representation is then fed into a softmax classifier, which outputs a
distribution over possible labels. The entire network, including the
embeddings, is trained using backpropagation.

The simpler version of the comparison concatenates the two
input vectors before feeding them into a standard neural network (NN)
layer.  The more powerful neural tensor network (NTN) version uses an
additional third-order tensor parameter to allow for multiplicative
interactions between the two inputs \cite{chen2013learning}. For more
details on the implementation and training of the layer functions, see
\cite{Bowman:Potts:Manning:2014}.

@@

This model used here differs from the one described in that work in
two ways. First, because the inputs are single terms, we do not use
a composition function here. Second, for our experiment on WordNet data, 
we introduce an additional neural network layer between the embedding 
input and the comparison function, which facilitates initializing the 
embeddings themselves from pretrained vectors and was found to help 
performance in that setting.

### 2014-10-12 writing/AAAI SSS '15/wordnet.tex
Edits

We extract three types of relation from WordNet. \ii{Hypernym} and
\ii{hyponym} can be are represented directly in the WordNet graph
structure, and correspond closely to the $\natrev$ and $\natfor$
relations from natural logic. As in natural logic, these relations are
mirror images of one another: if \ii{dog} is a hyponym of \ii{animal}
(perhaps indirectly by way of \ii{canid}, \ii{mammal}, etc.), then
\ii{animal} is a hypernym of \ii{dog}. We also extract \ii{coordinate}
terms, which share a direct hypernym, like \ii{dalmatian}, \ii{pug},
and \ii{puppy}, which are all direct hyponyms of \ii{dog}.  Coordinate
terms tend to exclude one another, thereby providing a loose approximation 
of the natural logic exclusion relation $\natneg$. 
WordNet defines hypernymy and hyponymy over
sets of synonyms, rather than over individual terms, so we do not
include a \ii{synonym} or \ii{equivalent} relation, but rather
consider only one member of each set of synonyms. Word pairs which do
not fall into these three relations are not included in the dataset.

To limit the size of the vocabulary without otherwise simplifying the learning problem, we extract all of the
instances of these three relations for single word nouns in WordNet that are hyponyms of the node 
\texttt{organism.n.01}. In order to balance the distribution of the classes, we slightly downsample instances 
of the \ii{coordinate} relation, yielding a total of 36,772 relations among 3,217 terms. We report results below using crossvalidation, choosing a disjoin 10\% test sample for each of five runs. Unlike in the previous experiment,
it is not straightforward here to determine in advance
how much data should be required to train an accurate model, so we performed training runs with 
various fractions of the remaining data. Embeddings were fixed at 25 dimensions and were initialized 
randomly or using distributional vectors from GloVe \cite{pennington2014glove}. The feature vector 
produced by the comparison layer was fixed at 80 dimensions.

\paragraph{Results} 
We find that the NTN performs perfectly with random initialization, and that the plain NN performs almost as well. We also find that initialization with GloVe is helpful in allowing the models to maintain fair performance
with smaller amounts of training data. Some of the randomly initialized model runs failed to learn
usable representations at all and labeled all examples with the most frequent labels. We excluded these runs from the statistics, but marked settings for which this occurred with the symbol $\dagger$. For all of the remaining runs, training accuracy was consistently above 99\%.

\begin{table}[h]
\centering\resizebox{5.5in}{!}{
  \setlength{\tabcolsep}{15pt}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabular}{r r@{ }r r@{ }r r@{ }r r@{ }r r@{ }r} 
    \toprule
     Portion of & \multicolumn{4}{c}{NN} & \multicolumn{4}{c}{NTN} & \multicolumn{2}{c}{Baseline}\\
     training data  & \multicolumn{2}{c}{w/ GloVe} & \multicolumn{2}{c}{w/o GloVe} & \multicolumn{2}{c}{w/ GloVe} & \multicolumn{2}{c}{w/o GloVe} \\
    \midrule
     100\%   & 99.7 &(0.0) & 99.4$^\dagger$ &(0.3)    & 99.6 &(0.0) & \textbf{100.0} &(0.1) & 37.9 &(--)\\
     33\%    & 96.0 &(0.5) & 95.3 &(0.2)     & 95.8 &(0.8)          & 95.5$^\dagger$ &(0.5) & 37.9 &(--)\\
     11\%    & 91.1 &(0.5) & 90.8$^\dagger$ &(0.3)    & 91.3 &(0.6)          & 90.9$^\dagger$ &(0.2) & 37.9 &(--)\\
     % 2\%    & 86.0 &(100.0) & ... &(100.0)   & 86.2 &(100.0)           & 76.4 &(87.3) & 37.9 &(--)\\ % Accidental run with slightly different train/test split.
    \bottomrule
  \end{tabular}}
 \caption{Mean test \% accuracy scores (with standard deviations) on the WordNet data over five-fold crossvalidation. The baseline figure is simply the frequency of the most frequent class, \ii{coordinate}.\label{b-table}} 
\end{table}

>>>>

We extract three types of relation from WordNet. \ii{Hypernym} and
\ii{hyponym} can be represented directly in the WordNet graph
structure, and correspond closely to the $\natrev$ and $\natfor$
relations from natural logic. As in natural logic, these relations are
mirror images of one another: if \ii{dog} is a hyponym of \ii{animal}
(perhaps indirectly by way of \ii{canid}, \ii{mammal}, etc.), then
\ii{animal} is a hypernym of \ii{dog}. We also extract \ii{coordinate}
terms, which share a direct hypernym, like \ii{dalmatian}, \ii{pug},
and \ii{puppy}, which are all direct hyponyms of \ii{dog}.  Coordinate
terms tend to exclude one another, thereby providing a loose approximation 
of the natural logic exclusion relation $\natalt$. 
WordNet defines hypernymy and hyponymy over
sets of synonyms, rather than over individual terms, so we do not
include a \ii{synonym} or \ii{equivalent} relation, but rather
consider only one member of each set of synonyms. Word pairs which do
not fall into these three relations are not included in the data set.

To limit the size of the vocabulary without otherwise simplifying the learning problem, we extract all of the
instances of these three relations for single word nouns in WordNet that are hyponyms of the node 
\texttt{organism.n.01}. In order to balance the distribution of the classes, we slightly downsample instances 
of the \ii{coordinate} relation, yielding a total of 36,772 relations among 3,217 terms. We report results below using crossvalidation, choosing a disjoint 10\% test sample for each of five runs. Unlike in the previous experiment,
it is not straightforward here to determine in advance
how much data should be required to train an accurate model, so we performed training runs with 
various fractions of the remaining data. Embeddings were fixed at 25 dimensions and were initialized 
randomly or using distributional vectors from GloVe \cite{pennington2014glove}. The feature vector 
produced by the comparison layer was fixed at 80 dimensions.

\paragraph{Results} 
We find that the NTN performs perfectly with random initialization, and that the plain NN performs almost as well,
a point of contrast with the results of Section~\ref{sec:join}. We also find that initialization with GloVe is helpful in allowing the models to maintain fair performance
with smaller amounts of training data. Some of the randomly initialized model runs failed to learn
usable representations at all and labeled all examples with the most frequent labels. We excluded these runs from the statistics, but marked settings for which this occurred with the symbol $\dagger$. For all of the remaining runs, training accuracy was consistently above 99\%.

\newcommand{\nodagger}{\phantom{$^\dagger$}}

\begin{table}[h]
\centering\resizebox{5.5in}{!}{
  \setlength{\tabcolsep}{15pt}
  \renewcommand{\arraystretch}{1.1}
  \begin{tabular}{r r@{ }r r@{ }r r@{ }r r@{ }r r@{ }r} 
    \toprule
     Portion of & \multicolumn{4}{c}{NN} & \multicolumn{4}{c}{NTN} & \multicolumn{2}{c}{Baseline}\\
     training data  & \multicolumn{2}{c}{w/ GloVe} & \multicolumn{2}{c}{w/o GloVe} & \multicolumn{2}{c}{w/ GloVe} & \multicolumn{2}{c}{w/o GloVe} \\
    \midrule
     100\%   & 99.7 &(0.0) & 99.4$^\dagger$ &(0.3)    & 99.6 &(0.0) & \textbf{100.0}\nodagger &(0.1) & 37.9 &(--)\\
     33\%    & 96.0 &(0.5) & 95.3\nodagger &(0.2)     & 95.8 &(0.8)          & 95.5$^\dagger$ &(0.5) & 37.9 &(--)\\
     11\%    & 91.1 &(0.5) & 90.8$^\dagger$ &(0.3)    & 91.3 &(0.6)          & 90.9$^\dagger$ &(0.2) & 37.9 &(--)\\
     % 2\%    & 86.0 &(100.0) & ... &(100.0)   & 86.2 &(100.0)           & 76.4 &(87.3) & 37.9 &(--)\\ % Accidental run with slightly different train/test split.
    \bottomrule
  \end{tabular}}
 \caption{Mean test \% accuracy scores (with standard deviations) on the WordNet data over five-fold crossvalidation. The baseline figure is the frequency of the most frequent class, \ii{coordinate}.\label{wordnet-table}} 
\end{table}
